---
title: "Refining infant research methods"
author: "Gonzalo García-Castro"
date: "27/9/2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(pwr)
library(tidyverse)
library(plotly)
library(RColorBrewer)
library(magrittr)
```

## Statistical power ($\beta$): a reminder

Probability of rejecting the null of hypothesis that there is no effect when there is an effect.
Provided that we want to achieve $\beta$ = .80, that Type I error rate ($\alpha$) = .05.

## Effect sizes in infant research

Effect sizes in infant looking-times reseach:

```{r oakes effect sizes, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
oakes <- read.delim("oakes.txt", row.names = NULL, header = FALSE)
colnames(oakes) <- c("citation", "n.cell", "test", "es", "power", "es50", "n.cell.needed", "n.cell.needed.es50")
oakeses <- oakes %>% select(es, es50) %>% gather(key = "estimation")
oakeses %$% estimation %>% factor
```

```{r oakes effect sizes plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(oakeses) +
  geom_histogram(aes(x = value, fill = estimation), binwidth = 0.2) +
  labs(x = "Estimated effect size", y = "Frequency") +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        text = element_text(size = 11)) +
  scale_fill_brewer(palette = "Pastel2")
```

# Increasing statistical power
Lakens


## Increase sample size (*n*)
Sample sizes in infant looking-time reseach:

```{r oakes sample sizes}
ggplot(oakes) +
  geom_histogram(aes(x = n.cell), binwidth = 5) +
  labs(x = "Subjects per cell", y = "Frequency") +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        text = element_text(size = 11)) +
  scale_fill_brewer(palette = "Pastel2")
```

- Collaborative approach

## Decreased measurement error

- Better measurements
- Reliability and inter-judge analysis

## Use within subject designs

## Use one sided-tests
Graph: Power analysis (one-sided vs. two-sided tests)

```{r data for power analysis, message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}
pwr.d5.data <- cbind(NULL, NULL)
for (i in 4:125){
  pwr.t.one.dep <- pwr.t.test(n = i, d = 0.5, type = "one.sample", alternative = "greater")
  pwr.t.two.dep <- pwr.t.test(n = i, d = 0.5, type = "one.sample", alternative = "two.sided")
  pwr.t.one.indep <- pwr.t.test(n = i, d = 0.5, type = "two.sample", alternative = "greater")
  pwr.t.two.indep <- pwr.t.test(n = i, d = 0.5, type = "two.sample", alternative = "two.sided")
  pwr.r.one <- pwr.r.test(n = i, r = 0.30, alternative = "greater")
  pwr.r.two <- pwr.r.test(n = i, r = 0.30, alternative = "two.sided")
  
  pwr.d5.data <- cbind(pwr.t.one.dep$n,
                       pwr.t.one.dep$power,
                       pwr.t.two.dep$power,
                       pwr.t.one.indep$power,
                       pwr.t.two.indep$power,
                       pwr.r.one$power,
                       pwr.r.two$power) %>% rbind(., pwr.d5.data) %>% as.data.frame
}

colnames(pwr.d5.data) <- c("n", "power.t.one.dep", "power.t.two.dep", "power.t.one.indep", "power.t.two.indep", "power.r.one", "power.r.two")
pwr.d5.data.long <- pwr.d5.data %>% select(power.t.one.dep,
                                           power.t.two.dep,
                                           power.t.one.indep,
                                           power.t.two.indep,
                                           power.r.one,
                                           power.r.two) %>% as.data.frame %>% gather(key = test, value = power) %>% add_column(n = rep(pwr.d5.data$n, 6))

pwr.d5.data.long$test %>% factor

n.d5.80.t.one.dep <- pwr.d5.data$n[abs(0.80-pwr.d5.data$power.t.one.dep) %>% which.min]
n.d5.80.t.two.dep <- pwr.d5.data$n[abs(0.80-pwr.d5.data$power.t.two.dep) %>% which.min]
n.d5.80.t.one.indep <- pwr.d5.data$n[abs(0.80-pwr.d5.data$power.t.one.indep) %>% which.min]
n.d5.80.t.two.indep <- pwr.d5.data$n[abs(0.80-pwr.d5.data$power.t.two.indep) %>% which.min]
n.d5.80.r.one <- pwr.d5.data$n[abs(0.80-pwr.d5.data$power.r.one) %>% which.min]
n.d5.80.r.two <- pwr.d5.data$n[abs(0.80-pwr.d5.data$power.r.two) %>% which.min]

n.d5.80 <- data.frame(n.d5.80.t.one.dep, n.d5.80.t.two.dep, n.d5.80.t.one.indep, n.d5.80.t.two.indep, n.d5.80.r.one, n.d5.80.r.two) %>% gather(key = test, value = n)

```

## When effect size is medium
```{r graph for power analysis medium, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(pwr.d5.data.long) +
  geom_line(aes(x = n, y = power, color = test), size = 1.5) +
  geom_hline(yintercept = 0.80, color = "grey") +
  geom_vline(data = n.d5.80, aes(xintercept = n), linetype = "dotted", size = 1, color = "grey") +
  geom_label(data = n.d5.80, aes(x = n, y = 0.80, label = paste("N =", n, sep = " ")), nudge_y = 0) +
  labs(title = "Sample size to reach 0.80 power, given a medium effect size", x = "Sample size", y = "Power") +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        legend.position = "bottom",
        text = element_text(size = 11)) +
  scale_color_brewer(palette = "BrBG")
```

## When effect size is small

```{r data for power analysis small, message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}
pwr.d2.data <- cbind(NULL, NULL)
for (i in 4:1000){
  pwr.t.one.dep <- pwr.t.test(n = i, d = 0.20, type = "one.sample", alternative = "greater")
  pwr.t.two.dep <- pwr.t.test(n = i, d = 0.20, type = "one.sample", alternative = "two.sided")
  pwr.t.one.indep <- pwr.t.test(n = i, d = 0.20, type = "two.sample", alternative = "greater")
  pwr.t.two.indep <- pwr.t.test(n = i, d = 0.20, type = "two.sample", alternative = "two.sided")
  pwr.r.one <- pwr.r.test(n = i, r = 0.10, alternative = "greater")
  pwr.r.two <- pwr.r.test(n = i, r = 0.10, alternative = "two.sided")
  
  pwr.d2.data <- cbind(pwr.t.one.dep$n,
                       pwr.t.one.dep$power,
                       pwr.t.two.dep$power,
                       pwr.t.one.indep$power,
                       pwr.t.two.indep$power,
                       pwr.r.one$power,
                       pwr.r.two$power) %>% rbind(., pwr.d2.data) %>% as.data.frame
}
colnames(pwr.d2.data) <- c("n", "power.t.one.dep", "power.t.two.dep", "power.t.one.indep", "power.t.two.indep", "power.r.one", "power.r.two")
pwr.d2.data.long <- pwr.d2.data %>% select(power.t.one.dep,
                                           power.t.two.dep,
                                           power.t.one.indep,
                                           power.t.two.indep,
                                           power.r.one,
                                           power.r.two) %>% as.data.frame %>% gather(key = test, value = power) %>% add_column(n = rep(pwr.d2.data$n, 6))

pwr.d2.data.long$test %>% factor

n.d2.80.t.one.dep <- pwr.d2.data$n[abs(0.80-pwr.d2.data$power.t.one.dep) %>% which.min]
n.d2.80.t.two.dep <- pwr.d2.data$n[abs(0.80-pwr.d2.data$power.t.two.dep) %>% which.min]
n.d2.80.t.one.indep <- pwr.d2.data$n[abs(0.80-pwr.d2.data$power.t.one.indep) %>% which.min]
n.d2.80.t.two.indep <- pwr.d2.data$n[abs(0.80-pwr.d2.data$power.t.two.indep) %>% which.min]
n.d2.80.r.one <- pwr.d2.data$n[abs(0.80-pwr.d2.data$power.r.one) %>% which.min]
n.d2.80.r.two <- pwr.d2.data$n[abs(0.80-pwr.d2.data$power.r.two) %>% which.min]

n.d2.80 <- data.frame(n.d2.80.t.one.dep, n.d2.80.t.two.dep, n.d2.80.t.one.indep, n.d2.80.t.two.indep, n.d2.80.r.one, n.d2.80.r.two) %>% gather(key = test, value = n)

```

```{r graph for power analysis, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(pwr.d2.data.long) +
  geom_line(aes(x = n, y = power, color = test), size = 1.5) +
  geom_hline(yintercept = 0.80, color = "grey") +
  geom_vline(data = n.d2.80, aes(xintercept = n), linetype = "dotted", size = 1, color = "grey") +
  geom_label(data = n.d2.80, aes(x = n, y = 0.80, label = paste("N =", n, sep = " ")), nudge_y = 0) +
  labs(title = "Sample size to reach 0.80 power, given a small effect size", x = "Sample size", y = "Power") +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        legend.position = "bottom",
        text = element_text(size = 11)) +
  scale_color_brewer(palette = "BrBG")
```

## Use continuous predictors (not categorical) when possible
The case of bilingualsism: monolinguals vs. bilinguals or degree of bilingualism.

## Increase variability in response variable
A 7-alternative item make it possible to find subtler differences than a 2-alternative item.

# Other ways of refining research
 
 * Pre-registration
 * Share data, code and materials
 * Replication (collaborative approach): Type I errors are corrected
    - ManyBabies
    - PsychAccelerator
    
    
 
 
 